Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-04-05 ms.tex
added scatter plot of EBV vs. tau
hugely improved presentation of two-component emission model

We adopt the thermal dust emission model of FDS99, which consists of two 
populations of dust grains emitting as MBBs, each with distinct 
emission/absorption characteristics. Importantly, the emissivity power 
law indices $\beta$ of the two species differ. The shape of this spectrum is 
given by:

@@

With $f_{1}$ = 0.0363, $\beta_1$ = 1.67, $\beta_2$ = 2.70, and $q_1/q_2$ = 
13.0. $\nu_0$ is chosen to be 3000 GHz. For the parameters thus far specified, 
the two dust temperatures are further related by:

\begin{equation}
T_1 = 0.352T_2^{1.18}
\end{equation}

@@

%the below statements seem like they might best be moved to a generic
%opening of the next section, stating our philosophy of breaking sky
%into pixels, each with 5 SED measurements, and that with this small
%number of constraints per pixel, it makes sense to have a baseline
% fit at full resolution that has only two free parameters
FDS99 found $T_2 \approx 16.2$K, $T_1 \approx 9.4$K to be typical at 
high-latitude. In the present formulation, there are only two free parameters 
to describe an observed SED: the hot dust temperature $T_2$ determines the SED 
shape and the normalization of $M_{\nu}$ determines the SED amplitude. In 
principle, we could allow additional parameters, such as $f_1$ and the ratio 
$q_1/q_2$, to vary. However the number of free parameters in our baseline fit 
at the full $6.1'$ resolution is limited by the fact that, for each sky 
location, the observed SED contains only five independent intensity 
measurements (\verb|i100| along with the four highest-frequency \PLANCK~bands).

@@

\section{Global Model Parameters}

\begin{deluxetable}{rrrrr} 
\tabletypesize{\scriptsize}
\tablecolumns{5} 
\tablewidth{0pc} 
\tablecaption{\label{obs} Global Model Parameters} 
\tablehead{
\colhead{Data} &
\colhead{$f_1$} &
\colhead{$q_1/q_2$} & 
\colhead{$\beta_1$} & 
\colhead{$\beta_2$}
}
\startdata
FIRAS+\IRAS & 0.0363 & 13.0 & 1.67 & 2.70 \\
\PLANCK+\IRAS & 0.0267 & 3.68 & 1.19 & 2.61
\enddata
\end{deluxetable}

@@

\subsection{Low-resolution Fits}
%mention that the pixelization changes ---> to lower HEALpix nside
%mention that over some region of the sky with high signal and high
%S/N we try fitting t2, normalization and one of q1/q2, f_1
%using more bands (including 143, and maybe 100 GHz)

@@

>>>>

% maybe a sentence to transition from preprocessing -> modeling
% e.g. With pre-processing completed, we turned to modeling the 
% thermal dust emission.

% mention that the two different emissivity power laws allow 
% the model to simultaneously explain narrow SED peak, and flattening
% of SED toward the millimeter
We adopt the thermal dust emission model of FDS99, which consists of two 
populations of dust grains emitting as MBBs, each with distinct 
emission/absorption characteristics. Importantly, the emissivity power law 
indices $\beta$ of the two species differ. The shape of this spectrum is given 
by:

@@

% With $f_{1}$ = 0.0363, $\beta_1$ = 1.67, $\beta_2$ = 2.70, and $q_1/q_2$ = 
% 13.0.

In accordance with the conventions of FDS99, $\nu_0$ is chosen to be 3000 GHz.
In principle, one could imagine this two MBB model to be specified by seven 
free parameters for every line of sight: $T_1$, $T_2$, $\beta_1$, $\beta_2$, 
$f_1$, and $q_1$/$q_2$ and the normalization of $M_{\nu}$. However, as 
discussed in FDS99, $T_1$=$T_1$($T_2$, $\beta_1$, $\beta_2$, $q_1/q_2$) is 
fully specified by these other parameters under the assumption that the 
temperature of each species is determined by maintaining thermal equilibrium 
with the same interstellar radiation field (ISRF). $T_1$ is always related to 
$T_2$ via a simple power law, although the prefactor and exponent depend on the
 parameters $q_1/q_2$, $\beta_1$ and $\beta_2$. $q_1/q_2$ encodes the
relative emission/absorption cross-sections of the two dust species, while
$f_1$ dictates the relative amounts of each species present.

These considerations still leave us with six potentially free parameters per 
line of sight. Unfortunately, fitting this many parameters per spatial pixel is
not feasible for our full-resolution $6.1$$'$ fits, as these are constrained by
only five broadband intensity measurements. Hence, as in FDS99, we 
deem certain parameters to be ``global'', i.e. spatially constant over the 
entire sky. In our full-resolution five-band fits, we designate $\beta_1$, 
$\beta_2$, $f_1$ and $q_1/q_2$ to be spatially constant. This same approach 
was employed by FDS99, and the globally best-fit values obtained by FDS99 for 
these parameters are listed in the first row of Table \ref{tab:global}. With 
these global parameters, FDS99 found $T_2$$\approx$$16.2$K, 
$T_1$$\approx$$9.4$K to be typical at high-latitude. In $\S$\ref{sec:global}, 
we discuss the best-fit global parameters favored by the \PLANCK~HFI data; 
these are listed in the second row of Table \ref{tab:global}.

Fixing the four parameters of Table \ref{tab:global}, our full-resolution, 
five-band fits have two remaining free parameters per line of sight: the hot 
dust temperature $T_2$ determines the SED shape and the normalization of 
$M_{\nu}$ determines the SED amplitude. In the lower-resolution fits of 
$\S$\ref{sec:lores} which include all HFI bands, we will allow $f_1$ to be a 
third free parameter, still holding $\beta_1$, $\beta_2$, and $q_1/q_2$ fixed.

%explain that we take the ``hot'' dust temperature to be T2 rather than T1

% maybe point to the SED fit summary figure as an example of what
%

% For the parameters thus far specified, the two dust temperatures are further 
% related by:

%\begin{equation}
%T_1 = 0.352T_2^{1.18}
%\end{equation}

@@

%the below statements seem like they might best be moved to a generic
%opening of the next section, stating our philosophy of breaking sky
%into pixels, each with 5 SED measurements, and that with this small
%number of constraints per pixel, it makes sense to have a baseline
% fit at full resolution that has only two free parameters

%FDS99 found $T_2 \approx 16.2$K, $T_1 \approx 9.4$K to be typical at 
%high-latitude. In the present formulation, there are only two free parameters 
%to describe an observed SED: the hot dust temperature $T_2$ determines the SED
%shape and the normalization of $M_{\nu}$ determines the SED amplitude. In 
%principle, we could allow additional parameters, such as $f_1$ and the ratio 
%$q_1/q_2$, to vary. However the number of free parameters in our baseline fit 
%at the full $6.1'$ resolution is limited by the fact that, for each sky 
%location, the observed SED contains only five independent intensity 
%measurements (\verb|i100| along with the four highest-frequency 
%\PLANCK~bands).

@@

\section{Global Model Parameters}
\label{sec:global}
%%%%

\begin{deluxetable}{rrrrr} 
\tabletypesize{\scriptsize}
\tablecolumns{5} 
\tablewidth{0pc} 
\tablecaption{\label{tab:global} Global Model Parameters} 
\tablehead{
\colhead{Data} &
\colhead{$f_1$} &
\colhead{$q_1/q_2$} & 
\colhead{$\beta_1$} & 
\colhead{$\beta_2$}
}
\startdata
FIRAS+\IRAS & 0.0363 & 13.0 & 1.67 & 2.70 \\
\PLANCK+\IRAS & 0.0267 & 3.68 & 1.19 & 2.61
\enddata
\end{deluxetable}

@@

\subsection{Low-resolution Fits}
\label{sec:lores}
%mention that the pixelization changes ---> to lower HEALpix nside
%mention that over some region of the sky with high signal and high
%S/N we try fitting t2, normalization and one of q1/q2, f_1
%using more bands (including 143, and maybe 100 GHz)

@@

\section{Calibrating to Optical Reddening}
\label{sec:ebv}

%FILL THIS SECTION IN !!!!!

% scatterplot figure showing linear regression of EBV vs. tau
\begin{figure}
\begin{center}
\epsfig{file=calib_ebv.eps, width=3.3in}
\caption{\label{fig:calib} Linear fit of $E(B-V)_{SSPP}$ as function of
$\tau_{545}$.}
\end{center}
\end{figure}

% need 2d histogram figure that is as closely analogous as possible to
% Figure 2 of Eddie's draft at : 
% http://faun.rc.fas.harvard.edu/eschlafly/2dmap/2dmap.pdf
